Cut systems and bicategories II

### 2015-08-29 cutsystems_cuts.tex
New version of cut systems paper.

\subsection{Splitting homotopies from connections}

>>>>

\section{Defining Cut systems}

To complete the construction of the cut system we need to finish defining the associator and unitors (a task begun at the outset of the section) and to prove that the coherence conditions hold.

Earlier the unitors were described by explicit formulas. Here we begin more abstractly: let $X \in \cat{L}(W,V)$ be a matrix factorisation with $W \in k[x_1,\ldots,x_n]$ and $V \in k[z_1,\ldots,z_m]$. Since $\Delta_V$ is the unit in $\LG$ there is a homotopy equivalence $\lambda: \Delta_V \otimes_{k[z]} X \lto X$ and so by Proposition \ref{prop:final_homotopy_equiv_cut} there are homotopy equivalences
\begin{equation}
\xymatrix@C+3pc{
\Delta_V \l X \ar@<-1ex>[r]_-{\Phi^{-1}} & S_m \otimes_k ( \Delta_V \otimes X )\ar@<-1ex>[l]_-{\Phi} \ar@<-1ex>[r]_-{1 \otimes \lambda} & S_m \otimes_k X \ar@<-1ex>[l]_-{1 \otimes \lambda^{-1}} 
}\,.
\end{equation}
By construction these isomorphisms are $A_m$-linear, and the map from right to left is an isomorphism $\mu: X \lto \Delta_V \l X$ in $\cat{L}(W,V)^\bullet$ corresponding under Lemma \ref{lemma:morphism_out_1} to
\[
\mu_0 = \xymatrix@C+1pc{
X \ar[r]^-{\lambda^{-1}} & \Delta_V \otimes_{k[z]} X \ar[r]^-{\varrho} & \Delta_V \l X
}\,.
\]
where $\varrho$ is the quotient map. Given the explicit formula for $\lambda^{-1}$ from \cite[Section 4.1]{lgdual}, this map $\mu_0$ has the explicit formula given earlier in \eqref{eq:explicit_mu0}. Since $\mu_0$ is natural in $X$, this also proves that $\mu$ is a natural isomorphism. The details for $\eta$ are similar, so the unitors are now defined.

The remaining components of the cut system $\cat{L}$ are the units $\Delta_W \in \cat{L}(W,W)$, which are the same as those for $\LG$ given in Section \ref{section:superbicatLG}, and the associators and unitors.

The unitors can also be described by explicit formulas. Given $X \in \cat{L}(W,V)$ as above with homogeneous basis $e_i$ as a $\mathbb{Z}_2$-graded $k[x,z]$-module, we define morphisms of matrix factorisations
\begin{gather}
\mu_0: X \lto \Delta_V \l X \,, \\
\mu_0(e_i) = \sum_{l \ge 0} \sum_{i_1 < \cdots < i_l} \sum_j \Theta_{i_1} \ldots \Theta_{i_l} \left\{ \partial^{z,z'}_{[i_l]}(d_X) \cdots \partial^{z,z'}_{[i_1]}(d_X) \right\}_{ji} \otimes e_j\label{eq:explicit_mu0}
\end{gather}
and
\begin{gather}
\eta_0: X \lto X \l \Delta_W \,,\\
\eta_0(e_i) = \sum_{l \ge 0} \sum_{i_1 < \cdots < i_l} \sum_j (-1)^{\binom{l}{2} + l|e_i|} e_j \otimes \left\{ \partial^{x,x'}_{[i_1]}(d_X) \cdots \partial^{x,x'}_{[i_l]}(d_X) \right\}_{ji} \Theta_{i_1} \ldots \Theta_{i_l}
\end{gather}
To make sense of these formulas, note that $\Delta_V$ is a matrix factorisation of $V(z) - V(z')$ over $k[z,z']$ and $\Delta_W$ is a matrix factorisation of $W(x) - W(x')$ over $k[x,x']$, while the divided difference operators are defined by
\[
{}^{t_i}(-): k[x,x'] \lto k[x,x']\,, \qquad f \mapsto f|_{x_i \mapsto x'_i}
\]
and
\[
\partial^{x,x'}_{[i]}: k[x,x'] \lto k[x,x']\,, \qquad f \mapsto \frac{{}^{t_1 \ldots t_{i-1}} f - {}^{t_1 \ldots t_i} f}{x_i - x'_i}
\]
and in the same way for $z,z'$. For a concrete example of these formulas see \cite{lgdual_survey}. The actual unitors are the morphisms $X \lto \Delta_V \l X$ and $X \lto X \l \Delta_W$ in $\cat{L}(W,V)^\bullet$ corresponding to $\mu_0$ and $\eta_0$ under the bijection of Lemma \ref{lemma:morphism_out_1}. It is not meant to be clear \emph{a priori} that these formulas even define morphisms, let alone that they satisfy the axioms for unitors; all these checks will be performed below.

\section{Convolution as a morphism of schemes}

We either fix a basis for Jacobians or assume $k$ a field so we can use Grobner.

\begin{itemize}
\item Restate hmf as matrices
\item Do idempotent completion as matrices with chosen homotopies
\item Do Clifford thickening as matrices with chosen homotopies
\item Write down the morphism of schemes
\end{itemize}

% This is the definition from cutsys10
\begin{definition} A \emph{cut system} $\cat{A}$ is determined by the following data:
\begin{itemize}
\item A class $|\cat{A}|$ of objects and for each object $a$ an integer $\pi(a) \ge 0$.
\item For each pair $a,b$ of objects a small idempotent complete supercategory $\cat{A}(a,b)$. 
\item For each triple $a,b,c$ of objects a cut functor of degree $\pi(b)$
\[
T: \cat{A}(a,b) \otimes_k \cat{A}(b,c) \lto \cat{A}(a,c)^{\bullet}\,.
\]
The image of a pair $X \otimes Y$ under the canonical extension
\[
\widetilde{T}: \cat{A}(a,b)^\bullet \otimes_k \cat{A}(b,c)^\bullet \lto \cat{A}(a,c)^\bullet
\]
is denoted $Y \l X$ and is called the \emph{cut of $Y$ and $X$ along $b$}.
\item For each object $a$, an object $\Delta_a \in \cat{A}(a,a)$.
\item For objects $a,b,c,d$ a natural isomorphism $\alpha$ as shown below:
\[
\xymatrix@C+2pc{
\cat{A}(a,b) \times \cat{A}(b,c) \times \cat{A}(c,d) \ar[d]_-{T \times 1} \ar[r]^-{1 \times T} & \cat{A}(a,b) \times \cat{A}(b,d)^\bullet \ar[d]^{\widetilde{T}} \ar@{=>}[dl]^-{\alpha}\\
\cat{A}(a,c)^\bullet \times \cat{A}(c,d) \ar[r]_-{\widetilde{T}} & \cat{A}(a,d)^\bullet
}
\]
That is, for $X \in \cat{A}(a,b), Y \in \cat{A}(b,c), Z \in \cat{A}(c,d)$ an isomorphism
\[
\alpha_{Z,Y,X}: ( Z \l Y ) \l X \lto Z \l ( Y \l X )
\]
of $A_{\pi(b)} \otimes_k A_{\pi(c)}$-modules in $\cat{A}(a,d)$, natural in all three variables.
\item For objects $a,b$ natural isomorphisms $\mu, \eta$ as shown below:
\[
\xymatrix@C+2pc{
1 \times \cat{A}(a,b) \ar[d]_-{\iota_{\Delta} \times 1} & \cat{A}(a,b) \ar@{=>}@<4ex>[l]_-{\eta} \ar@{=>}@<-4ex>[r]^-{\mu} \ar[d]^-{\iota} \ar[l]_-{\cong} \ar[r]^-{\cong} & \cat{A}(a,b) \times 1 \ar[d]^-{1 \times \iota_{\Delta}}\\
\cat{A}(a,a) \times \cat{A}(a,b) \ar[r]_-{T} & \cat{A}(A,B)^\bullet & \cat{A}(a,b) \times \cat{A}(b,b) \ar[l]^-{T}
}
\]
where $1$ is the category with one object and one morphism and $\iota_{\Delta}$ is the functor sending this object to $\Delta$. That is, for each $X \in \cat{A}(a,b)$ a pair of isomorphisms
\begin{equation}\label{eq:defn_cutsys_lambda}
\mu: X \lto \Delta_b \l X, \quad \eta: X \lto X \l \Delta_a
\end{equation}
in $\cat{A}(a,b)^\bullet$ natural in $X$.
\end{itemize}
This data is subject to the following coherence conditions:
\begin{itemize}
\item[(1)] For objects $a,b,c,d,e$ and $X \in \cat{A}(a,b),Y \in \cat{A}(b,c),Z \in \cat{A}(c,d),Q \in \cat{A}(d,e)$, commutativity of the diagram
\begin{equation}
\xymatrix@C+2pc{
((Q \l Z) \l Y ) \l X \ar[d]_-{\alpha_{Q \l Z, Y, X}}\ar[r]^-{\alpha_{Q,Z,Y} \l 1} & (Q \l (Z \l Y)) \l X \ar[r]^-{\alpha_{Q, Z \l Y, X}} & Q \l ( ( Z \l Y ) \l X ) \ar[d]^-{1 \l \alpha_{Z,Y,X}}\\
(Q \l Z) \l (Y \l X) \ar[rr]_-{\alpha_{Q, Z, Y \l X}} & & Q \l ( Z \l ( Y \l X ) )
}
\end{equation}
Note that when we write $\alpha_{Q, Z \l Y, X}$ we refer only the underlying object of $Z \l Y$.
\item[(2)] For objects $a,b$ and $X \in \cat{A}(a,b), Y \in \cat{A}(b,c)$ commutativity of the diagram
\begin{equation}\label{eq:unitor_constraint_cut}
\xymatrix{
(Y \l \Delta_b ) \l X \ar[rr]^-{\alpha} & & Y \l ( \Delta_b \l X )\\
& Y \l X \ar[ul]^-{\eta \l 1_X} \ar[ur]_-{1_Y \l \mu}
}
\end{equation}
\item[(3)] commutativity of \eqref{eq:psiunitor1},\eqref{eq:psiunitor2} for every $X$ and \eqref{eq:psithirdlastdia}, \eqref{eq:psimiddledia}, \eqref{eq:psilastdia} for every composable triple $X,Y,Z$.
\end{itemize}
%Finally, the data is subject to three computability conditions:
%\begin{itemize}
%\item[(1)] For each pair of objects $a,b$ the category $\cat{A}(a,b)$ is effectively computable, in the sense that the morphisms and objects are described by finite data and the composition rule is computable.
%\item[(2)] For each triple $a,b,c$ of objects the associated cut functor is effectively computable. In particular, for $1$-morphisms $Y$ and $X$ the object $Y \l X$ and its Clifford action are effectively computable from $X$ and $Y$.
%\item[(3)] The associators and unitors are effectively computable from their arguments.
%\end{itemize}
\end{definition}

Let $\cat{A}$ be a cut system. There is a superbicategory $\cat{A}^\bullet$ with objects the class of objects of $\cat{A}$ and with $1$- and $2$-morphisms defined by $\cat{A}^\bullet(a,b) = \cat{A}(a,b)^\bullet$. The units of $\cat{A}^\bullet$ are the objects $\Delta_a$ in the cut system, and the composition in the bicategory is given by the canonical extensions $\widetilde{T}$. To complete the definition of a superbicategory it only remains to give the associators and unitors, and the super-structure.

\begin{lemma} Given objects $a,b,c,d$ the natural isomorphism $\alpha$ canonically extends to a natural isomorphism $\widetilde{\alpha}$ as in the diagram
\[
\xymatrix@C+2pc{
\cat{A}(a,b)^\bullet \times \cat{A}(b,c)^\bullet \times \cat{A}(c,d)^\bullet \ar[d]_-{\widetilde{T} \times 1} \ar[r]^-{1 \times \widetilde{T}} & \cat{A}(a,b)^\bullet \times \cat{A}(b,d)^\bullet \ar[d]^{\widetilde{T}} \ar@{=>}[dl]^-{\widetilde{\alpha}}\\
\cat{A}(a,c)^\bullet \times \cat{A}(c,d)^\bullet \ar[r]_-{\widetilde{T}} & \cat{A}(a,d)^\bullet
}
\]
\end{lemma}
\begin{proof}
Suppose $(X,n) \in \cat{A}(a,b)^\bullet, (Y,m) \in \cat{A}(b,c)^\bullet$ and $(Z,t) \in \cat{A}(c,d)^\bullet$ are given. The underlying object of $((Z,t) \l (Y, m)) \l (X,n)$ is $(Z \l Y) \l X$, and it is clear that $\alpha_{Z,Y,X}$ is $A_t \otimes A_m \otimes A_n$-linear and therefore defines an isomorphism in $\cat{A}(a,d)^\bullet$
\[
\alpha_{Z,Y,X}: ((Z,t) \l (Y, m)) \l (X,n) \lto (Z,t) \l ((Y, m) \l (X,n))\,.
\]
Naturality is straightforward.
\end{proof}

\begin{lemma} Given objects $a,b$ the natural isomorphisms $\eta, \mu$ canonically extend to natural isomorphisms $\widetilde{\eta}, \widetilde{\mu}$ as in the diagram
\[
\xymatrix@C+2pc{
1 \times \cat{A}(a,b)^\bullet \ar[d]_-{\iota_{\Delta} \times 1} & \cat{A}(a,b)^\bullet \ar@{=>}@<4ex>[l]_-{\widetilde{\eta}} \ar@{=>}@<-4ex>[r]^-{\widetilde{\mu}} \ar[d]^-{1} \ar[l]_-{\cong} \ar[r]^-{\cong} & \cat{A}(a,b)^\bullet \times 1 \ar[d]^-{1 \times \iota_{\Delta}}\\
\cat{A}(a,a)^\bullet \times \cat{A}(a,b)^\bullet \ar[r]_-{\widetilde{T}} & \cat{A}(A,B)^\bullet & \cat{A}(a,b)^\bullet \times \cat{A}(b,b)^\bullet \ar[l]^-{\widetilde{T}}
}
\]
 \end{lemma}
\begin{proof}
Let $(X,n) \in \cat{A}(a,b)^\bullet$. Forgetting the $A_n$-action on $X$, we have isomorphisms $\eta$ and $\mu$ as in \eqref{eq:unitor_tilde_pre}, which are respectively $A_{\pi(b)}$ and $A_{\pi(a)}$-linear. Since these isomorphisms are natural, they are automatically $A_n$-linear. Verifying that these isomorphisms are natural with respect to the larger class of morphisms in $\cat{A}(a,b)^\bullet$ is straightforward.
\end{proof}

The supercategory structure is given as part of the cut system, so we have:

% The above definition of a cut system is from cutsys10. This is by p.11 there something which gives rise to a cut system in the sense of cutsys6, which is proved there to define a superbicategory.
\begin{proposition} Given a cut system $\cat{A}$, $\cat{A}^\bullet$ is a superbicategory.
\end{proposition}
\begin{proof}
This follows from Lemma \ref{lemma:constructingsuper}, with associator $\widetilde{\alpha}$ and unitors $\lambda =\widetilde{\mu}^{-1}, \rho = \widetilde{\eta}^{-1}$. The coherence follows from the coherence axioms of the cut system.
\end{proof}

We often confuse a cut system $\cat{A}$ with the associated superbicategory. For example, we refer to objects $X$ of $\cat{A}(a,b)^\bullet$ as a $1$-morphisms. To distinguish objects of the original category $\cat{A}(a,b)$ we refer to these objects as \emph{normal} or \emph{cut-free} $1$-morphisms. We keep the notation $Y \l X$ for the composition rule in $\cat{A}^\bullet$.

The following lemma is useful in verifying the unitor constraint.

\begin{lemma}\label{lemma:unitor_check_help} The commutativity of \eqref{eq:unitor_constraint_cut} is equivalent to commutativity of
\begin{equation}\label{eq:unitor_check_help1}
\xymatrix@C+1pc{
(Y \l \Delta_b) \l X \ar[rr]^-{\alpha} & & Y \l ( \Delta_b \l X )\\
Y \l X \ar[u]^-{\eta_0 \l 1} & Y \l X \ar[l]^-{e_{\pi(b)}} \ar[r]_-{e_{\pi(b)}} & Y \l X \ar[u]_-{1 \l \mu_0}
}
\end{equation}
where $e_{\pi(b)}$ is the idempotent of Definition \ref{defn:idempotent_e}.
\end{lemma}
\begin{proof}
Let $m = \pi(b)$. Then $(Y \l \Delta_b) \l X$ has two $A_m$ actions, one from $Y \l \Delta_b$ which we denote by $A_{m_1}$, and the other from the cut with $X$ which we denote by $A_{m_2}$. By definition of the canonical extension $\widetilde{T}$ in Appendix \ref{app:proof_cut_extension}, $\eta \l 1$ is the $A_{m_1+m_2}$-linear morphism
\[
\eta: S_{m_1 + m_2, m_2} \otimes_{A_{m_2}} ( Y \l X ) \lto (Y \l \Delta_b ) \l X
\]
which corresponds under Lemma \ref{lemma:morphisms_two_forms} to
\[
\xymatrix@C+1pc{
Y \l X \ar[r]^-{e_m} & Y \l X \ar[r]^-{\eta_0 \l 1} & (Y \l \Delta_b) \l X
}\,.
\]
Applying the same logic to $\mu$ completes the proof.
\end{proof}

In a cut system the unitors $\mu, \eta$ go in the opposite direction to what one would expect from the conventions of bicategories. This is because it is more natural to describe morphisms out of a cut-free $1$-morphism and into a $1$-morphism carrying a nontrivial Clifford representation, than in the other direction.

\begin{proof}[Proof of Theorem \ref{theorem:l_cut_system}] We first check the coherence axioms. The first and third are trivial. To prove (2) we need by Lemma \ref{lemma:unitor_check_help} only check commutativity of \eqref{eq:unitor_check_help1} in the usual situation of $X: W \lto V$ and $Y: V \lto U$.

This amounts to checking commutativity of
\begin{equation}\label{eq:coherence3_1}
\xymatrix{
(Y \l \Delta_V ) \l X \ar[rr]^-{\alpha} & & Y \l ( \Delta_V \l X )\\
(Y \otimes \Delta_V) \otimes J_V \otimes X \ar[u]^-{\varrho \otimes 1} & & Y \otimes J_V \otimes ( \Delta_V \otimes X) \ar[u]_-{1 \otimes \varrho}\\
Y \otimes J_V \otimes X \ar[u]^-{ \rho^{-1} \otimes 1} & Y \otimes J_V \otimes X \ar[l]^-{e_m} \ar[r]_-{e_m} & Y \otimes J_V \otimes X \ar[u]_-{1 \otimes \lambda^{-1}}
}\,.
\end{equation}
By definition $e_m$ factors as
\[
\xymatrix{
Y \l X \ar[r]^-{\Phi^{-1}} & S_m \otimes_k ( Y \otimes X ) \ar[r] & Y \otimes X \ar[r] & S_m \otimes_k (Y \otimes X) \ar[r]^-{\Phi} & Y \l X
}
\]
so commutativity of \eqref{eq:coherence3_1} follows from commutativity of
\[
\xymatrix@C-0.5pc{
(Y \l \Delta_V ) \l X \ar[rrr]^-{\alpha} & & & Y \l ( \Delta_V \l X )\\
(Y \otimes \Delta_V) \otimes J_V \otimes X \ar[u]^-{\varrho \otimes 1} & (Y \otimes \Delta_V ) \otimes X \ar[l]_-{\varrho} \ar[r]^-{\alpha} & Y \otimes (\Delta_V \otimes X) \ar[r]^-{\varrho} & Y \otimes J_V \otimes ( \Delta_V \otimes X) \ar[u]_-{1 \otimes \varrho}\\
Y \otimes J_V \otimes X \ar[u]^-{ \rho^{-1} \otimes 1} & Y \otimes X \ar[u]^-{\rho^{-1} \otimes 1} \ar@{=}[r] \ar[l]^-{\varrho} & Y \otimes X \ar[u]_-{1 \otimes \lambda^{-1}} \ar[r]_-{\varrho} & Y \otimes J_V \otimes X \ar[u]_-{1 \otimes \lambda^{-1}}
}\,.
\]
The innermost square of this diagram is the coherence constraint for the unitors in $\LG$.

%Finally, we check the computability conditions. The associator and unitors are derived explicitly from $d_X$ and $d_Y$, so (3) is clear, and (2) is addressed by Remark \ref{remark:inflation} and the explicit formulas for the Clifford action. For (1), the objects and morphisms of $\cat{L}(W,V)$ are described by finite matrices of polynomials, and the composition rule is given by multiplication of matrices. There is one subtlety: morphisms are matrices of polynomials \emph{identified up to homotopy}.

%To argue that $\cat{L}(W,V)$ is effectively computable we need to show that it is possible to determine algorithmically whether two morphisms are homotopic. This is important even at the level of objects: an object is a pair $(X,e)$ where $e^2 = e$ holds up to homotopy. Without an effective homotopy relation, there would be no algorithmic way to even distinguish between objects and non-objects. The necessary check is in Section \ref{example:computing_homs}.
\end{proof}

We will use \emph{effectively computable} as a synonym for the precise notion of a \emph{computable} function \cite{enderton}, since ``computable'' often has a more colloquial meaning in our context.

\begin{remark} To give natural isomorphisms \eqref{eq:defn_cutsys_lambda} in $\cat{A}(a,b)^\bullet$ is equivalent to giving natural isomorphisms in $\cat{A}(a,b)$
\begin{equation}\label{eq:unitor_tilde_pre}
\mu: S_{\pi(b)} \otimes_k X \lto \Delta_b \l X, \quad \eta: S_{\pi(a)} \otimes_k X \lto X \l \Delta_a
\end{equation}
which are respectively $A_{\pi(b)}$ and $A_{\pi(a)}$-linear.

By Lemma \ref{lemma:morphism_out_1} such morphisms correspond bijectively to morphisms in $\cat{A}(a,b)$
\begin{equation}\label{eq:unitor_tilde}
\mu_0: X \lto \Delta_b \l X, \quad \eta_0: X \lto X \l \Delta_a
\end{equation}
satisfying respectively $a_i^\dagger \circ \mu_0 = 0$ for $1 \le i \le m$ and $a_i^\dagger \circ \eta_0 = 0$ for $1 \le i \le n$. The reason to introduce these maps is that commutativity of \eqref{eq:unitor_constraint_cut} can be more conveniently expressed in terms of the maps in \eqref{eq:unitor_tilde}, see Lemma \ref{lemma:unitor_check_help} below.
\end{remark}

\subsection{Cut models}\label{section:cut_models}

\begin{definition} A \emph{cut model} for a superbicategory $\cat{B}$ is a cut system $\cat{A}$ with the same underlying class of objects as $\cat{B}$ and a strong superfunctor $I: \cat{B} \lto \cat{A}^\bullet$ with the property that for every pair of objects $a,b$ the functor
\[
I_{a,b}: \cat{B}(a,b) \lto \cat{A}(a,b)^\bullet
\]
factors as
\[
\xymatrix{
\cat{B}(a,b) \ar[rr]^-{I_{a,b}} \ar[dr]_-{J_{a,b}} & & \cat{A}(a,b)^\bullet\\
& \cat{A}(a,b) \ar[ur]_-{\iota}
}
\]
where $J_{a,b}$ is an equivalence of supercategories. We often identify $X$ with $I(X)$.
\end{definition}

A cut model is an effectively computable model for composition of $1$-morphisms in~$\cat{B}$. Given $X: a \lto b$ and $Y: b \lto c$, let us suppose that the composition $Y \circ X$ in $\cat{B}$ is not defined algorithmically: perhaps this object is only identified up to unique isomorphism by a universal property. In contrast, the cut $Y \l X$ is computable from $X$ and $Y$, and the strong superfunctor $I$ provides an isomorphism in $\cat{A}(a,c)^\bullet$
\begin{equation}
Y \l X \cong Y \circ X\,.
\end{equation}
Here $Y \circ X$ is embedded into $\cat{A}^\bullet$ as a cut-free $1$-morphism and the cut $Y \l X$ by definition comes with an $A_{\pi(b)}$-module structure. That is, there is an isomorphism of $A_{\pi(b)}$-modules
\begin{equation}\label{eq:cutvsoriginal}
Y \l X \cong S_{\pi(b)} \otimes_k (Y \circ X )
\end{equation}
or what is the same, an isomorphism of objects in $\cat{A}(a,b)$
\begin{equation}\label{eq:cutvsoriginal2}
S^*_{\pi(b)} \otimes_{A_{\pi(b)}} ( Y \l X ) \cong Y \circ X \,.
\end{equation}
The general philosophy is that a cut model decomposes composition in $\cat{B}$ into two stages: 
\begin{itemize}
\item first form the cut $Y \l X$, and then
\item tensor $Y \l X$ with $S^*_{\pi(b)}$ to compute $Y \circ X$. 
\end{itemize}
The idea is that the desired composition $Y \circ X$ lies inside $Y \l X$ in a way that is ``known'' in the sense that it is encoded by the Clifford action, but which is only \emph{realised} or \emph{constructed} after a computation is done. Forming the tensor product $S^*_{\pi(b)} \otimes_{A_{\pi(b)}} ( Y \l X )$ starting with the concretely known object $Y \l X$ can be further decomposed into a $\pi(b)$-step process of ``atomic'' computations, using the diagram of Remark \ref{remark:tensor_and_idempotent}
\begin{equation}\label{eq:computation_fiber_diagram}
\xymatrix@C+1pc@R+1pc{
Z_0 \cong Y \circ X \ar@{.}[d] & Z_1 \ar@{.}[d] & & Z_{\pi(b)-1} \ar@{.}[d] & Z_{\pi(b)} = Y \l X \ar@{.}[d]\\
0 \ar@{.>}@<0.8ex>[r] & 1 \ar@{.>}@<0.8ex>[r]\ar@{.>}@<0.8ex>[l] & \cdots \ar@{.>}@<0.8ex>[l]\ar@{.>}@<0.8ex>[r] & \pi(b)-1  \ar@{.>}@<0.8ex>[l]\ar@{.>}@<0.8ex>[r] & \pi(b) \ar@{.>}@<0.8ex>[l]
}\,.
\end{equation}
% remark:tensor_in_steps.
The intermediate objects are $Z_i = S_{i, \pi(b)} \otimes_{A_{\pi(b)}} ( Y \l X )$ for $i < \pi(b)$. Each of these objects $Z_i$ is obtained from $Z_{i+1}$ by splitting an idempotent of the form $a^\dagger a$.

Suppose we are in the situation where $Y \circ X$ is identified by a universal property but an explicit object realising this universal property is not given as part of $\cat{B}$. Then the result of the $\pi(b)$-step computation depicted above is an explicit object $Z_0$ of $\cat{A}(a,c) \cong \cat{B}(a,c)$ with this universal property. Thus the cut system gives a process for constructing representing objects for the uniquely defined isomorphism class $Y \circ X$ of $\cat{B}(a,c)$.

Conceptually, the number of steps in \eqref{eq:computation_fiber_diagram} represents the computational cost of a cut across the object $b$. For any cut across this object, $\pi(b)$-steps are necessary to convert the ``uncertainty'' represented by $Y \l X$ with its $A_{\pi(b)}$-action to the certainty of an object of $\cat{A}(a,c)$ (in the sense of Remark \ref{remark:tensor_and_idempotent}). In the example of the next section, the cost $\pi(W)$ of cutting across a hypersurface singularity $W$ is determined by its dimension.

\section{Cut systems}\label{section:cut_systems}

We begin with the definition of cut functors.

\begin{definition} Given supercategories $\cat{C}_1,\cat{C}_2,\cat{C}_3$ a \emph{superfunctor} $T: \cat{C}_1 \otimes_k \cat{C}_2 \lto \cat{C}_3$ is a functor together with natural isomorphisms
\begin{align*}
\tau &: T \circ ( \Psi \otimes 1 ) \lto \Psi \circ T,\\
\tau &: T \circ ( 1 \otimes \Psi ) \lto \Psi \circ T
\end{align*}
making the following diagrams commute:
\begin{equation}\label{eq:psicomp1}
\xymatrix@R+1pc@C+1pc{
T( 1 \otimes \Psi^2 ) \ar[r]^-{\tau}\ar[d]_-{T(1 \otimes \xi)} & \Psi T( 1 \otimes \Psi ) \ar[r]^-{\tau} & \Psi^2 T \ar[d]^-{\xi * 1_T}\\
T \ar[rr]_-{1_T} & & T
}
\end{equation}
\begin{equation}\label{eq:psicomp2}
\xymatrix@R+1pc@C+1pc{
T( \Psi^2 \otimes 1 ) \ar[r]^-{\tau} \ar[d]_-{T (\psi \otimes 1)}  & \Psi T( \Psi \otimes 1 ) \ar[r]^-{\tau} & \Psi^2 T \ar[d]^-{\xi * 1_T}\\
T \ar[rr]_-{1_T} & & T
}
\end{equation}
\begin{equation}\label{eq:psicomp3}
\xymatrix@R+1pc@C+1pc{
T( \Psi \otimes \Psi ) \ar[r]^-{\tau}\ar[d]_-{-\tau} & \Psi T( 1 \otimes \Psi ) \ar[d]^-{\tau}\\
\Psi T( \Psi \otimes 1 ) \ar[r]_-{\tau} & \Psi^2 T
}
\end{equation}

\end{definition}

\begin{definition}\label{defn:cutfunctor1} A \emph{cut functor of degree $t$} on a triple $\cat{C}_1, \cat{C}_2,\cat{C}_3$ of idempotent complete small supercategories is a superfunctor $T: \cat{C}_1 \otimes_k \cat{C}_2 \lto \cat{C}_3^\bullet$ such that
\[
\xymatrix{
\cat{C}_1 \times \cat{C}_2 \ar[r]^-{T} & \cat{C}_3^\bullet \ar[r]^-{p} & \mathbb{N}
}
\]
sends every object to the integer $t \in \mathbb{N}$.
\end{definition}

With the above notation:

\begin{proposition}\label{prop:extend_cut_functor} Given a cut functor $T$ of degree $t$ on $\cat{C}_1, \cat{C}_2,\cat{C}_3$ there is a canonical superfunctor $\widetilde{T}: \cat{C}_1^\bullet \otimes_k \cat{C}_2^\bullet \lto \cat{C}_3^\bullet$ with the property that the diagram
\[
\xymatrix{
\cat{C}_1 \times \cat{C}_2 \ar[dr]_-{\iota \times \iota} \ar[rr]^-{T} & & \cat{C}_3^\bullet\\
& \cat{C}_1^\bullet \times \cat{C}_2^\bullet \ar[ur]_-{\widetilde{T}}
}
\]
commutes, and for every pair of objects $X \in \cat{C}_1^\bullet$ and $Y \in \cat{C}_2^\bullet$
\[
p \widetilde{T}( X, Y ) = p( X ) + p( Y ) + t\,.
\]
We refer to $\widetilde{T}$ as the \emph{extension} of $T$.
%and
%\[
%\xymatrix@C+2pc{
%\cat{C}_1^\bullet \times \cat{C}_2^\bullet \ar[d]_-{p \times p}\ar[rr]^-{\widetilde{T}} & & \cat{C}_3^\bullet \ar[dd]^-{p}\\
%\mathbb{N} \times \mathbb{N} \ar[d]_-{\cong}\\
%\mathbb{N} \times \{ * \} \times \mathbb{N} \ar[r]_-{1 \times t \times 1} & \mathbb{N} \times \mathbb{N} \times \mathbb{N} \ar[r]_-{\otimes} & \mathbb{N}
%}
%\]
%commute. 
\end{proposition}
\begin{proof}
The proof is given in Appendix \ref{app:proof_cut_extension}.
\end{proof}

\section{Extending cut functors}\label{app:proof_cut_extension}

The purpose of this appendix is to give the proof of Proposition \ref{prop:extend_cut_functor}. First, note that

\begin{lemma}\label{lemma:altcomppersp} Let $\cat{C}$ be a small idempotent complete supercategory and take morphisms
\[
\xymatrix{
(X,n) \ar[r]^-{\alpha} & (Y,m) \ar[r]^-{\beta} & (Z,l)
}\\
\]
in $\cat{C}^\bullet$. Under \eqref{eq:isodefatensor} these correspond to bimodule maps
\[
\widetilde{\alpha}: S_{m,n} \lto \cat{C}^*(X,Y)\,, \qquad \widetilde{\beta}: S_{l,m} \lto \cat{C}^*(Y,Z)\,.
\]
The composite $\beta \circ \alpha: S_{l,n} \otimes_{A_n} X \lto Z$ in $\cat{C}^\bullet$ corresponds to the composite
\[
\xymatrix{
S_{l,n} \cong S_{l,m} \otimes_{A_m} S_{m,n} \ar[r]^-{\widetilde{\beta} \otimes \widetilde{\alpha}} & \cat{C}^*(Y,Z) \otimes_{A_m} \cat{C}^*(X,Y) \ar[r]^-{\circ} & \cat{C}^*(X,Z)
}\,.
\]
\end{lemma}
\begin{proof}
Follows by evaluation on $1 \otimes 1^* \in S_{l,n}$.
\end{proof}

\begin{proof}[Proof of Proposition \ref{prop:extend_cut_functor}]
Given $(X,n) \in \cat{C}_1^\bullet$ and $(Y,m) \in \cat{C}_2^\bullet$ we have $T(X,Y) = (Z,t)$ for some object $Z$ of $\cat{C}_3$. Given $a \in A_n$, say of odd degree, there is a morphism of $A_t$-modules $Z \lto \Psi Z$ given by
\[
\xymatrix@C+2pc{
T(X,Y) \ar[r]^-{T(a \otimes 1)} & T(\Psi X, Y ) \ar[r]^-{\tau} & \Psi T(X,Y)
}
\]
and this defines an algebra morphism $A_n \lto \cat{C}_3^*(Z,Z)$. Doing the same for $A_m$ defines a morphism $A_{m+t+n} \cong A_m \otimes A_t \otimes A_n \lto \cat{C}_3^*(Z,Z)$ and an object $(Z, m + t +n) \in \cat{C}_3^\bullet$. We define $\widetilde{T}$ on objects by
\begin{equation}\label{eq:ttilde}
\widetilde{T}( (X,n), (Y,m) ) = (Z, m + t + n)\,.
\end{equation}
Next we define $\widetilde{T}$ on morphisms. Given $\alpha: (X_1,n_1) \lto (X_2,n_2)$ in $\cat{C}_1^\bullet$ and $T(X_i,Y) = (Z_i,t)$ for $i = 1,2$ we must construct a morphism
\[
\widetilde{T}(\alpha,1): (Z_1, m+t+n_1) \lto (Z_2, m+t+n_2)\,.
\]
Since $\alpha$ corresponds to an $A_{n_1}$-linear map $\widetilde{\alpha}: V \lto \cat{C}_1^*(X_1,X_2)$ we can compose with $T$ to get an $A_{n_2}$-$A_{n_1}$-bilinear map $\eta := T(1,Y) \circ \widetilde{\alpha}: V \lto \cat{C}_3^*(Z_1,Z_2)$. This extends to a morphism of $A_{n_2+c}$-$A_{n_1+c}$-bimodules
\[
V \otimes_k A_c \lto \cat{C}_3^*(Z_1,Z_2), \qquad v \otimes a \mapsto \eta(v) a\,.
\]
There is a canonical isomorphism of $A_{n+2+c}$-$A_{n_1+c}$-bimodules
\begin{align*}
S_{n_2+c,n_1+c} \cong S_{n_2+c} \otimes_k S_{n_1+c}^* \cong (S_{n_2} \otimes_k S_c) \otimes_k (S_c^* \otimes_k S_{n_1}^*) \cong V \otimes_k (S_c \otimes_k S_c^*) \cong V \otimes_k A_c
\end{align*}
so we have constructed $S_{n_2+c,n_1+c} \lto \cat{C}_3^*(Z_1,Z_2)$ which corresponds to a morphism of $A_{n_2+c}$-modules $S_{n_2+c, n_1+c} \otimes_{A_{n_1+c}} Z_1 \lto Z_2$ which we define to be $\widetilde{T}(\alpha,1)$. Functoriality of this construction follows from Lemma \ref{lemma:altcomppersp}. The functoriality in the other variable follows similarly. By construction $\widetilde{T}$ makes both of the required diagrams commute.
\end{proof}

\subsection{The cut model of $\LG$}
